---
title: Interactive Maps with R
author: Alex Mandel, Nistara Randhawa, Ryan Peek, Mahesh Maskey
date: January 30, 2019
output: github_document
#always_allow_html: yes
---

# Introduction

* What is an Interactive Map? 
    * A map that allows you to pan, zoom, and query(click) in the plot without additional code, or plotting. Also known as Slippy Maps, Dynamic Maps, etc...
* When should I use an Interactive Map? 
    * Data and result exploration.
* When should I not use an Interactive Map? 
    * When your map will be viewed in anything other than a web browser. 
    * When you don't need the viewer to be able to pan, zoom, and click on things in the map.



```{r map, echo=FALSE, fig.cap="Static Map", message=FALSE}
# Static Map vs Dynamic Map
library(mapview)
data(breweries, package="mapview")

plot(breweries$geometry, col = "purple")
#mapview(breweries)	
```

[Interactive Map of the same data](https://r-spatial.github.io/mapview/articles/articles/mapview_01-basics.html)


## Goal

Give people an dive into how to make interactive plots in R with maps. Focus on maps that allow the user to pan, zoom, query(click).

### Core tools

Mapview & Leaflet Packages

Mapview wraps Leaflet, making it simpler to use. You can use anything from leaflet as additional arguments to Mapview.

# Tutorial

## Exercise 1
Mapview [Article 1](https://r-spatial.github.io/mapview/articles/articles/mapview_01-basics.html)

Note: the background may not show up in Rstudio, but if you open the file in browser it will. 

## Exercise 2
Mapview [Article 2](https://r-spatial.github.io/mapview/articles/articles/mapview_02-advanced.html)

### Basemaps

If you want to use a different Basemap (Background Map), here's a list of easy to use examples. http://leaflet-extras.github.io/leaflet-providers/preview/index.html
You can use any other XYZ tile service, read the documentation on how.

## Other Tips

### Labels

There's a bunch of different methods for labels, [read more](https://rdrr.io/cran/mapview/man/addStaticLabels.html)


```{r labels, fig.cap="Labelling Examples"}
library(leaflet)
library(mapview)

# Turn on labels to start
lopt = labelOptions(noHide = TRUE,
                    direction = 'top',
                    textOnly = TRUE)


# Point labels
m1 = mapview(breweries)
l1 = addStaticLabels(m1,
                     label = breweries$brewery,
                     labelOptions = lopt)
#l1

# Polygon Labels
m2 = mapview(franconia)
l2 = addStaticLabels(m2,
                     label = franconia$NAME_ASCI,
                     labelOptions = lopt)
#l2

# Custom labels
m3 = m2 + m1
l3 = addStaticLabels(m3,
                     data = franconia,
                     label = franconia$NAME_ASCI,
                     labelOptions = lopt)
#l3
```

### Scale Bar
Adding a Scale Bar to the Map
```{r measure, fig.cap="Scale Bar Example"}
# Take an existing map object and add a scale bar
m2 <- addMeasure(l3)
#m2
```

### Saving your map
You can also save the interactive map to an HTML file which you can share with others. 
```{r savemap, eval = FALSE}
m2 = mapview(franconia)
mapshot(m2, "franconia.html")
```

If you used `leaflet` for your interactive map, you would use `saveWidget` from the `htmlwidgets` package. 
```{r savemap2, eval = FALSE}
library(leaflet)
library(htmlwidgets)
data(quakes)

# Map first 20 rows from the `quakes` dataset
quake_map = leaflet(data = quakes[1:20,]) %>%
    addTiles() %>%
    addMarkers(~long, ~lat, popup = ~as.character(mag))

saveWidget(quake_map, "quakes.html")
```


# Outline

## Data Types 
- Points, Lines, Polygons, Rasters
- Vectors include the option for popups from attribute tables
- sp and sf

## Data Examples
- Crop Field Boundaries (polygons) with UAS RGB and NDVI rasters (From Taylors presentation last quarter), it could be crop type
- Getting country/administrative boundaries (polygons) via R and using them in Leaflet. Useful when working with health data or any field project in different countries
- Climate data

## Steps 
- Data Preparation
- Decide how to derive data
- Dealing with missing data
- Compile them at a point wise

## Styling 
- Points - circles vs markers, their attributes, using custom icons, etc
- Lines
- Polygons
- Rasters
- Basemaps
- Standard - OSM, Bing, Google, 
- XYZ tiles - i.e. Stamen, etc
- Transparency
- Labels
- Attribute pop-ups
- Toggle different data layers on the map
- Tools that allow measurement/point extraction (Javascript based but easy to drop code in to play with (Leaflet))

## Bonus
Exporting to share with others (HTML or Shiny)

## References
- https://rstudio.github.io/leaflet/
- https://r-spatial.github.io/mapview/
