---
title: Interactive Maps with R
author: Alex Mandel, Nistara Randhawa, Ryan Peek, Mahesh Maskey
date: January 30, 2019
output: github_document
#always_allow_html: yes
---

# Introduction

* What is an Interactive Map? 
    * A map that allows you to pan, zoom, and query(click) in the plot without additional code, or plotting. Also known as Slippy Maps, Dynamic Maps, etc...
* When should I use an Interactive Map? 
    * Data and result exploration.
* When should I not use an Interactive Map? 
    * When your map will be viewed in anything other than a web browser. 
    * When you don't need the viewer to be able to pan, zoom, and click on things in the map.
    * When the data is really big.
* What could I use instead?
    * Other static plot methods in R
    * Desktop GIS applications like QGIS and ArcGIS (Handles large data much better)



This is a Static Map:

```{r map, echo=FALSE, fig.width=4 , fig.cap="Static Map", message=FALSE}
# Static Map vs Dynamic Map
library(mapview)
data(breweries, package="mapview")

plot(breweries$geometry, col = "purple")
#mapview(breweries)	
```

For comparison see this [interactive map of the same data](https://r-spatial.github.io/mapview/articles/articles/mapview_01-basics.html)


## Goal

Give people an dive into how to make interactive plots in R with maps. Focus on maps that allow the user to pan, zoom, query(click).

### Core tools

[Mapview](https://r-spatial.github.io/mapview/) & [Leaflet](https://rstudio.github.io/leaflet/) Packages

Mapview wraps Leaflet, making it simpler to use. You can use anything from leaflet as additional arguments to Mapview.

# Tutorial

## Exercise 1
Mapview [Article 1](https://r-spatial.github.io/mapview/articles/articles/mapview_01-basics.html)

Note: the backgrounds may not show up in Rstudio, but if you open the file in a web browser (upper right corner of the RStudio preview window) it will. 

## Exercise 2
Mapview [Article 2](https://r-spatial.github.io/mapview/articles/articles/mapview_02-advanced.html)

### Basemaps

If you want to use a different Basemap (Background Map), [here's a list of easy to use maps](http://leaflet-extras.github.io/leaflet-providers/preview/index.html).
You can use any other XYZ tile services, read the [documentation on addTiles](https://www.rdocumentation.org/packages/leaflet/versions/2.0.2/topics/addControl) for various options.

*Note* Some tiles source(e.g. Google), require that you have an api key to use their basemaps.

## Other Tips

### Labels

There's a bunch of different methods for labels, [read more](https://rdrr.io/cran/mapview/man/addStaticLabels.html)


```{r labels, eval=FALSE, fig.cap="Labelling Examples"}
library(leaflet)
library(mapview)

# Turn on labels to start
lopt = labelOptions(noHide = TRUE,
                    direction = 'top',
                    textOnly = TRUE)


# Point labels
m1 = mapview(breweries)
l1 = addStaticLabels(m1,
                     label = breweries$brewery,
                     labelOptions = lopt)

# Call the mapview to plot it
l1

# Polygon Labels
m2 = mapview(franconia)
l2 = addStaticLabels(m2,
                     label = franconia$NAME_ASCI,
                     labelOptions = lopt)
l2

# Custom labels
m3 = m2 + m1
l3 = addStaticLabels(m3,
                     data = franconia,
                     label = franconia$NAME_ASCI,
                     labelOptions = lopt)
l3
```

### Colors and Legends

Depending on the data type (raster, vector, etc) being plotted, colors can be modified using either a name or [hexcode](https://htmlcolorcodes.com/). 

```{r colors, eval=FALSE}

mapview(franconia, 
        color="gray", # some colors work as color
        col.regions="#008080") # more refined

```

We can also assign a color palette based on the data:

```{r zcolors, eval=FALSE}

# assign color based on variable name in data:
(zcol_map1 <- mapview(franconia, zcol="district"))

# assign color but change the NA value and legend title
mapview(breweries, zcol="number.of.types", na.color="transparent",
        layer.name="Number of Beer Types <br> at Each Brewery")

# manipulate some data to plot on top:
library(dplyr)
smallbrews <- breweries %>% filter(number.of.types<2) # only make 1 beer!
manybrews <- breweries %>% filter(number.of.types>8) # make more than 8 kinds of beer!

# and layer with beer data:
zcol_map1 + 
  mapview(smallbrews, color="white", col.regions="black",
        na.color="transparent", 
        layer.name="Single Beer Breweries") +
  mapview(manybrews, color="white", col.regions="maroon",
          na.color="transparent", 
          layer.name="Multi Beer Breweries")

```

### Scale Bar
Adding a Scale Bar to the Map
```{r measure, eval=FALSE, fig.cap="Scale Bar Example"}
# Take an existing map object and add a scale bar
m2 <- addMeasure(l3)
m2
```

### Saving your map
You can also save the interactive map to an HTML file which you can share with others. 
```{r savemap, eval = FALSE}
m2 = mapview(franconia)
mapshot(m2, "franconia.html")
```

If you used `leaflet` for your interactive map, you would use `saveWidget` from the `htmlwidgets` package. 
```{r savemap2, eval = FALSE}
library(leaflet)
library(htmlwidgets)
data(quakes)

# Map first 20 rows from the `quakes` dataset
quake_map = leaflet(data = quakes[1:20,]) %>%
    addTiles() %>%
    addMarkers(~long, ~lat, popup = ~as.character(mag))

saveWidget(quake_map, "quakes.html")
```


# Things that are possible

## Data Types 

* Points, Lines, Polygons, Rasters
* Vectors include the option for popups from attribute tables
* sp and sf

## Styling 

* Points - circles vs markers, their attributes, using custom icons, etc
* Lines
* Polygons
* Rasters
* Basemaps
    * Standard - OSM, Bing, Google, 
    * XYZ tiles - i.e. Stamen, etc
* Transparency
* Labels
* Attribute pop-ups
* Toggle different data layers on the map
* Tools that allow measurement/point extraction (Javascript based but easy to drop code in to play with (Leaflet))

## Bonus

* Exporting to share with mapshot.
* More advanced, you can use Leaflet in Shiny apps for 2 way interaction with users (they can draw on the maps and you can use that as data) 